The perceptual dimensions of \\ sonority-driven epenthesis

### 2013-09-30 paper.tex
Converted Irish examples to textipa, added conclusion to machinery section

\ex. Detailed comparison of {\sc Sonority Rise} vs {\sc Sonority Angle} hierarchies:

@@

Syllable Contact applies only across syllable boundaries.  Therefore, when sonority-driven epenthesis occurs in onsets or codas, other sonority-based markedness constraints must be employed.  One such is  Sonority Sequencing \citep{selkirk.1984}, which states that sonority must be strictly increasing in the onset and strictly decreasing in the coda. \citep{rose.2000} uses two variants of this constraint, one strict and one non-strict, in her analysis of Chaha.

@@

\ex. \a. /alb\textipa{@}/ $\rightarrow$ [al\textipa{@b@}] \\
         {\it Alba} `Scotland'
     \b. /dorx\textipa{@}/ $\rightarrow$ [dor\textipa{@x@}] \\
         {\it dorcha} `dark'
     \b. /gorm/  $\rightarrow$  [gor\textipa{@}m] \\
         {\it gorm} `blue'
     \b. /banb\textipa{@}/ $\rightarrow$ [ban\textipa{@b@}] \\
         {\it Banba} `a name for Ireland'
     \b. /kon\textipa{f@}/ $\rightarrow$ [kon\textipa{@f@}] \\
         {\it confadh} `anger'
     \b. /an\textsuperscript{j}m\textsuperscript{j}/ $\rightarrow$ [an\textsuperscript{j}im\textsuperscript{j}] \\
         {\it ainm} `name'
     \z.
     \citep[(37)]{carnie.1994} and \citep[(2)]{ni.chiosain.1999}

@@

\ex. \a. /farno\textipa{:g}/ $\rightarrow$ [farno\textipa{:g}], *[far\textipa{@}no\textipa{:g}] \\
         {\it fearnog} `alder'
     \b. /bord/ $\rightarrow$ [bord], *[bor\textipa{@}d] \\
         {\it bord} `table'
     \b. /kalr\textipa{@}/ $\rightarrow$ [kalr\textipa{@}], *[kal\textipa{@r@}] \\
         {\it calra} `calorie'  
     \z. 
     \citep[(52)]{carnie.1994}\footnote{\citet{ni.chiosain.1999} has a more complicated set of facts as to which homorganic clusters are broken up by epenthesis. As this is not the focus of this paper, I will assume the simpler dataset that Carnie presents.}

@@

\ex. \a. /kork/ $\rightarrow$ [kork], *[kor\textipa{@}k] \\
         {\it Cork} `Cork (place name)' 
     \b. /kork\textipa{@}/ $\rightarrow$ [kork\textipa{@}], *[kor\textipa{@k@}] \\
         {\it corca} `people' 
     \b. /korp/  $\rightarrow$ [korp], *[kor\textipa{@p}] \\
         {\it corp} `body' 
     \b. /\textipa{baNk}/ $\rightarrow$ [\textipa{baNk}], *[\textipa{baN@k}] \\
         {\it banc} `bank'
     \z.
     \citep[(39a,b,c,26c)]{carnie.1994}

>>>>

\ex. Detailed comparison of {\sc Sonority Angle} vs {\sc Sonority Rise} hierarchies:

@@

Compared with the cluster hierarchy predicted by {\sc Sonority Angle}, we again see several differences.

\ex. {\sc Sonority Angle} hierarchy, repeated for purposes of comparison:

I will not enumerate all the differences in predictions, but notice two crucial distinctions: the {\sc Dis}-based Syllable Contact hierarchy predicts equal susceptibility to epenthesis of all the clusters with a {\sc Dis} of +1 (RN, NF and FT), and thus it does not distinguish RN from the remainder of the falling sonority clusters. In addition, if NT and RT fail to undergo epenthesis, this predicts that RF should not undergo epenthesis either. Both of these are contrary to the empirical evidence of Chaha and Irish.

Another difference between the markedness-based account and the sonority contour faithfulness approach is that Syllable Contact applies only across syllable boundaries, while the {\sc Ident(Son$\measuredangle$)}$<$n constraints are agnostic with regard to syllable structure. When sonority-driven epenthesis occurs elsewhere, for example in codas, other sonority-based markedness constraints must be employed.  One such is Sonority Sequencing \citep{selkirk.1984}, which states that sonority must be strictly increasing in the onset and strictly decreasing in the coda. In both the Irish and Chaha cases, we see sonority-driven epenthesis occurring both in heterosyllabic and coda cluster contexts, and we will see that adopting the family of {\sc Ident(Son$\measuredangle$}$<$n constraints economises on the number of constraints we need to refer to sonority.

In this section, I have introduced all the necessary machinery to analyse cases of sonority-driven epenthesis in Irish and Chaha, as well as to understand the competing analyses. We now proceed to the case study of Irish.

@@

\ex. \a. /\textipa{alb@}/ $\rightarrow$ [\textipa{al@b@}] \\
         {\it Alba} `Scotland'
     \b. /\textipa{dorx@}/ $\rightarrow$ [\textipa{dor@x@}] \\
         {\it dorcha} `dark'
     \b. /\textipa{gorm}/  $\rightarrow$  [\textipa{gor@m}] \\
         {\it gorm} `blue'
     \b. /\textipa{banb@}/ $\rightarrow$ [\textipa{ban@b@}] \\
         {\it Banba} `a name for Ireland'
     \b. /\textipa{konf@}/ $\rightarrow$ [\textipa{kon@f@}] \\
         {\it confadh} `anger'
     \b. /\textipa{an\textsuperscript{j}m\textsuperscript{j}}/ $\rightarrow$ [\textipa{an\textsuperscript{j}im\textsuperscript{j}}] \\
         {\it ainm} `name'
     \z.
     \citep[(37)]{carnie.1994} and \citep[(2)]{ni.chiosain.1999}

@@

\ex. \a. /\textipa{farno:g}/ $\rightarrow$ [\textipa{farno:g}], *[\textipa{far@no:g}] \\
         {\it fearnog} `alder'
     \b. /\textipa{bord}/ $\rightarrow$ [\textipa{bord}], *[\textipa{bor@d}] \\
         {\it bord} `table'
     \b. /\textipa{kalr@}/ $\rightarrow$ [\textipa{kalr@}], *[\textipa{kal@r@}] \\
         {\it calra} `calorie'  
     \z. 
     \citep[(52)]{carnie.1994}\footnote{\citet{ni.chiosain.1999} has a more complicated set of facts as to which homorganic clusters are broken up by epenthesis. As this is not the focus of this paper, I will assume the simpler dataset that Carnie presents.}

@@

\ex. \a. /\textipa{kork}/ $\rightarrow$ [\textipa{kork}], *[\textipa{kor@k}] \\
         {\it Cork} `Cork (place name)' 
     \b. /\textipa{kork@}/ $\rightarrow$ [\textipa{kork@}], *[\textipa{kor@k@}] \\
         {\it corca} `people' 
     \b. /\textipa{korp}/  $\rightarrow$ [\textipa{korp}], *[\textipa{kor@p}] \\
         {\it corp} `body' 
     \b. /\textipa{baNk}/ $\rightarrow$ [\textipa{baNk}], *[\textipa{baN@k}] \\
         {\it banc} `bank'
     \z.
     \citep[(39a,b,c,26c)]{carnie.1994}
